---
title: "2018 Final Exam"
author: "Salinas, Alexa"
date: "12/17/2019"
output: word_document
---

## 1. Write a regular expression, and only the regular expression, for each of the three descriptions below:

a. a DNA sequence beginning with AAATTT and containing 10-15 codons (triplets of DNA bases),
including the initial AAATTT

AAATTT([ATCG]{3}){10,15}$

b. any of the words cot, rot, hot, shot, short, fort

[crhf]o[rt]

## 2. The directory home contains three directories (records, poems, photos). Both records and poems contain
a set of .txt files that you will use for the tasks described below:

a. Assuming you begin in the home directory on a Unix system, provide the Unix code required to make
a new directory archive in the records directory, copy all .txt files from records to archive, and
delete any files with a filename starting with ‘old’ from the records directory.

cd records #moving into records directory
mkdir archive #making a new directory in the records directory called archive
cp *.txt archive/ #copying any files that end in .txt to the directory archive
rm old* #removing any files in the records directory that start with old

b. Assuming you begin in the home directory on a Unix system, provide the Unix code required to append
your initials to the beginning of the filename for all .txt files in the poems directory. Use a for-loop to
accomplish this task.

cd poems #moving into poems directory*
for file in *.txt #for the file in the set of files that end in .txt
do
cat $file > JW$file #cat the file into a file with the appended name
done

## 3. The information below is contained in the file lunch.txt. Provide Unix code needed to generate a new file (Danslunch.txt) that contains only entries for dan that are sorted by menu item and changes all occurrences
of dan to Dan.
dan hamburger
alice tacos
dan waffles
dan salad
ralph pizza
alice burrito

cat lunch.txt | grep dan | sort -k 2 | sed ‘s/dan/Dan/g’ > Danslunch.txt

- concatenates the file, finds the lines with dan in it, sorts the second column (food) alphabetically, substitutes the word dan for Dan, outputt it into new file

## 4. Earlier in the semester you coded a guessing game choosing an integer between 1 and 100. A non-functional version of R code for this task is below and provided in a script called guessing_game.R. Fix any errors in the code and adjust the code so that the player only has 5 attempts to guess the correct answer or they lose.

# randomly draw an integer between 1 and 100
myNumber=sample(x=1:50,size=1)
# create a while loop that runs until we indicate the guesser is correct
correct=0
while(correct==0){
# prompt the user for a guess
guess=readline(promp="Guess:")
if(guess<theNumber)
  print("Higher")
else{
  print("Lower")
}
if(guess=myNumber){
  print("Correct")
  correct=1
  }
}

```{r 4A.}

# randomly draw an integer between 1 and 100
myNumber=sample(x=1:100,size=1) #drawing a number btwn 1 and 100, not 1 and 50

# create a while loop that runs until we indicate the guesser is correct
correct=0 #starting the number of guesses (correct)at 0
while(correct<5){ #while the number of guesses (correct) is less than 5
  #prompt the user for a guess
  guess=readline(prompt = "Guess:") #prompts user to put in number to guess
  
  if(guess<myNumber){ #if the guess is less than the number
    print("Higher") #prints higher
    correct=correct+1 #adds 1 to the number of guesses, so correct increases by 1
  }else{
    if (guess>myNumber){ #if guess is greater than the number
      print("Lower") #prints lower
    correct=correct+1 #adds 1 to the number of guesses, so correct increases by 1
    }
  }
  if(guess==myNumber){ #if guess equals the number
    print("Correct") #print "correct"
  }
}

#once correct=5, it ends the game and you are not allowed to guess anymore

```

5. Write a function in R that takes a vector as an argument and returns the sum of even numbers contained
in the vector. Remember %% will return the remainder of integer division. Also be sure to write your function to handle the case where the vector supplied contains no even numbers.

```{r 5A.}

sum_even_num<-function(vector){ #creating a function for a vector that sums the even numbers
  evensum=0 #starting even sum at 0
  for (i in 1:length(vector)){ # for observation in the vector
    if (vector[i]%%2==0){ #if the remainder when the value is divided by 2 equals zero (an even number)
      evensum=evensum+vector[i] #add the value to evensum
    }
  }
return(evensum) #returns even sum
}

#usage with a vector 
x=(1:10)
sum_even_num(x)
#returns 30
#if no even numbers in vector, returns 0, because there is no sum of the even numbers

```

6. Assume that y has an even number of entries and that the first half of y come from Group 1 and the
second half of y come from Group 2. These data are available in y.csv. Provide the R code to generate a
barplot of the means of y for Group 1 and Group 2.

```{r 6A.}

y=read.csv(file="y.csv", header = TRUE) #loading the csv file
y$group=c(rep("Group 1",length(y)/2),rep("Group 2",length(y)/2))

ggplot(data = y) + 
geom_bar(aes(x=group, y=y), stat = "summary", fun.y="mean")+theme_classic()+xlab("Group")+ylab("mean of y")

```
































