Biocomp Notes Final Review 

These things will not be on the final: bioinformatics (muscle/hmmr),modeling,git,while loops,maps
get organized with files on laptop and know where everything is 
ICB - Practice Final Exam

#Nt+1=Nt
#Mt+1=Mt

1. You have a text file names "apples.txt" with the following content:
Apples are the best fruit.
They are the apple of my eye.
Apples grow on apple trees.
Would you like an apple?

Write unix code to replace all instances of "Apple" or "apple" with "orange".

1A.

*cat apples.txt | sed 's/[aA]pple/orange/'*

or

sed -e 's/apple/orange/g' -e 's/Apple/orange/g' apples.txt

*above codes replaces multiple words in one line of code using -e 

2. The following R code is intended to count how many values contained in the vector x are above 10 and
how many are below 10 and then print both values. Unfortunately, the code isn’t working. Fix the code so
that it works as intended:

count=1
count2=1
x<-c(5,8,11,13,15) #vector 5 values 
for(i in 1:length(count)){
  if(x[count]>10){
    count[i]=count+1
  }else(x[count2]<10)
    count2[i]=count2+1
}

print(c("There are this many values below 10 in x:",count))
print(c("There are this many values below 10 in x:",count2))

```{r 2A.}

x<-c(5,8,11,13,15) #setting vector w values
count=1
count2=1
for(i in 1:length(x)){
  if(x[count]>10){
    count=count+1
  }else if(x[count]<10){
    count2=count2+1
  }
}

print(c("There are this many values below 10 in x:",count2))
print(c("There are this many values above 10 in x:",count))

```

3. Write a regular expression for each of the three descriptions below:

a) any date after December 31st, 1999 in the format "01/01/2000".

*[0-9]{2}/[0-9]{2}/2[0-9]{3}*

b) a DNA sequence beginning with ATG adn containing 10-15 codons (triplets of A, T, C, or G)

*ATG([ATCG]{3}){10,15}*

c) any of the words raft, draft, graft, or shaft.

*aft is a simple one; a more specific one might be [dg?][rh]aft*

4. What is returned by the last line of this series of Unix commands?

$ pwd
/Users/stuart/
$ ls
a.txt b.txt c.txt d.txt
$ mkdir temp
$ mv *.txt ./temp
$ cp ./temp/a.txt ./
$ rm -r ./temp
$ ls

4A. 

*a.txt*

5. Write the R code required to plot the values contained in a vector y vs the values in a vectory x and add a spline best fit line to the scatter plot.

```{r 5A.}

x<-c(1,2,3,4,5)
y<-c(10,20,30,40,50)
library(ggplot2)
toplot<-data.frame(x=x,y=y)
ggplot(data=toplot,aes(x=x,y=y))+geom_point()+geom_smooth(method="loess")

```


6. What is the difference between a matrix and a dataframe in R?

6A. 

*A matrix can only hold one data mode, but a dataframe can hold multiple data modes.*






#Notes from last lec (12/11)
[a-zA-Z] #will match any 1 letter upper/lowercase 
#know character classes and how to quantify them
#grep -E
#sed -E 's/find/replace/g'

#1A on practice exam:
#AAATTT then 10-15 codons (3 of ATC or G); match AAATTT
([ATCG]{3}){8,13}

#Unix practice problem
for i in $@
  do
cat $i | grep -E "Stuart">>file.txt

#Unix shell script practice problem
#usage: bash script.sh filename 15
cat $1 | head $2 #$1 refers to filename while $2 refers to 15
#remember #@ refers to all arguments, not just 1st or 2nd 

#usage: bash script.sh  *.csv
#for exam: $1,$2,$@

#functions in R 
colin<-function(x,number){
value<-mean(x)/number
return(value)
}
colin(x=vector,fileout="stuart.csv")
#== means does it equal while != means does it not equal

#fasta file in Unix practice problem
#fasta files look like this
#this is called DNA.fasta
>seq01_05-01-2009
  ATCG
>seq02_10-10-2009
  ATTG

cat DNA.fasta | grep -E ">"



